InteroceptionScale - Data Analysis

### 2024-11-21 study1/analysis/2_analysis.qmd
script

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(EGAnet)
```

@@

```{r}
ega <- bootEGA(
  items,
  model = "glasso",  # BGGM
  algorithm = "louvain",  # walktrap
  EGA.type = "hierEGA",
  plot.itemStability=FALSE,
  typicalStructure=TRUE,
  plot.typicalGraph=TRUE,
  iter=100,
  seed=3, ncores = 4)

# EGAnet::dimensionStability(ega)
itemstability <- EGAnet::itemStability(ega, IS.plot=FALSE)
plot(itemstability)

itemstability_table <- itemstability$lower_order$item.stability$all.dimensions |>
  as.data.frame() |>
  datawizard::data_addprefix("C")
itemstability_table$Cluster <- colnames(itemstability_table)[max.col(itemstability_table, ties.method='first')]

itemstability_table <- itemstability_table |>
  rownames_to_column(var="Item") |>
  rowwise() |>
  mutate(Stability = max(c_across(-c(Item, Cluster)))) |>
  arrange(Cluster, desc(Stability))

toremove <- filter(itemstability_table, Stability < 0.5)$Item

itemstability_table |>
  select(-Cluster, -Stability) |>
  gt::gt() |>
  gt::data_color(columns=-Item,
                 method = "numeric",
                 palette = c("white", "orange", "green"),
                 domain = c(0, 1)) |>
  gt::data_color(columns=Item, fn=\(x) ifelse(x %in% toremove, "#FFEBEE", "white")) |>
  gt::tab_header(
    title = gt::md("**Item Stability**"),
    subtitle = "Proportion of times an item is assigned to the same cluster"
  )
```

## Exploratory Graph Analysis (EGA)

```{r}
ega <- bootEGA(
  items,
  model = "glasso",  # BGGM
  algorithm = "louvain",  # walktrap
  EGA.type = "hierEGA",
  # plot.EGA = FALSE,
  plot.itemStability=FALSE,
  typicalStructure=TRUE,
  plot.typicalGraph=TRUE,
  seed=3, ncores = 4)

plot(ega)
summary(ega)

# EGAnet::dimensionStability(ega)
itemstability <- EGAnet::itemStability(ega)
itemstability
plot(itemstability)

# Loadings
loadings <- ega$typicalGraph
loadings$parameters$lower_loadings
loadings$parameters$lower_scores
```

<!-- ```{r} -->
<!-- library(ggraph) -->

<!-- EGAnet::convert2tidygraph(ega$EGA) -->

<!-- graph <- ega$typicalGraph -->

<!-- edges_lower <- graph$lower_order$graph |>  -->
<!--   as.data.frame() |>  -->
<!--   rownames_to_column(var="from") |>  -->
<!--   pivot_longer(-from, names_to="to", values_to="weight") |>  -->
<!--   filter(weight > 0)  -->
<!-- nodes_lower <- graph$lower_order$typical.dim.variables -->

<!-- tidygraph::as_tbl_graph(edges_lower, nodes=nodes_lower, directed=FALSE) |>  -->
<!--   ggraph::ggraph(layout = 'circle') + -->
<!--   ggraph::geom_node_point() + -->
<!--   ggraph::geom_edge_bend(aes(edge_width = weight), alpha=0.5) + -->
<!--   # ggraph::geom_node_text(aes(label=item), repel=TRUE) + -->
<!--   theme_void() -->
<!-- ``` -->

@@

>>>>

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(EGAnet)
library(tidygraph)
library(ggraph)
```

@@

```{r}
#| eval: false

ega1 <- bootEGA(
  items,
  EGA.type = "EGA",
  model = "glasso",  # BGGM
  algorithm = "louvain",  # walktrap
  type="resampling",
  plot.itemStability=FALSE,
  typicalStructure=TRUE,
  plot.typicalGraph=TRUE,
  iter=500,
  seed=3, ncores = 4)

save(ega1, file="models/ega1.RData")
```

```{r}
load("models/ega1.RData")

# EGAnet::dimensionStability(ega)
itemstability <- EGAnet::itemStability(ega1, IS.plot=FALSE)
plot(itemstability)

make_loadingtable <- function(x) {
  t <- as.data.frame(x) |>
    datawizard::data_addprefix("C")
  t$Cluster <- colnames(t)[max.col(t, ties.method='first')]

  t |>
    rownames_to_column(var="Item") |>
    rowwise() |>
    mutate(Max = max(c_across(-c(Item, Cluster)))) |>
    arrange(Cluster, desc(Max)) |> 
    as.data.frame()
}

itemstability_table <- make_loadingtable(itemstability$item.stability$all.dimensions)

toremove2 <- filter(itemstability_table, Max < 0.7)$Item

itemstability_table |>
  select(-Cluster, -Max) |>
  gt::gt() |>
  gt::data_color(columns=-Item,
                 method = "numeric",
                 palette = c("white", "orange", "green"),
                 domain = c(0, 1)) |>
  gt::data_color(columns=Item, fn=\(x) ifelse(x %in% toremove2, "#FFEBEE", "white")) |>
  gt::tab_header(
    title = gt::md("**Item Stability**"),
    subtitle = "Proportion of times an item is assigned to the same cluster"
  )
```

```{r}
items <- select(items, -all_of(toremove2))
byfacet <- select(byfacet, -all_of(toremove2))
bymodality <- select(bymodality, -all_of(toremove2))
byrandom <- select(byrandom, -all_of(toremove2))
```

### Item Loadings

```{r}
#| eval: false

ega2 <- bootEGA(
  items,
  EGA.type = "hierEGA",
  model = "glasso",  # BGGM
  algorithm = "louvain",  # walktrap
  type="resampling",
  plot.itemStability=FALSE,
  typicalStructure=TRUE,
  plot.typicalGraph=TRUE,
  iter=500,
  seed=3, ncores = 4)

save(ega2, file="models/ega2.RData")
```

```{r}
load("models/ega2.RData")

# itemStability(ega2)

lower <- make_loadingtable(net.loads(ega2$EGA$lower_order)$std)

toremove3 <- lower$Item[!lower$Item %in% slice_max(lower, Max, n=3, by="Cluster")$Item]

lower |>
  select(-Cluster, -Max) |>
  gt::gt() |>
  gt::data_color(columns=-Item,
                 method = "numeric",
                 palette = c("red", "white", "green"),
                 domain = c(-1, 1)) |>
  gt::data_color(columns=Item, fn=\(x) ifelse(x %in% toremove3, "#FFEBEE", "white")) |>
  gt::tab_header(
    title = gt::md("**Item Loadings**"),
    subtitle = "Node centrality"
  )
```

```{r}
items <- select(items, -all_of(toremove3))
byfacet <- select(byfacet, -all_of(toremove3))
bymodality <- select(bymodality, -all_of(toremove3))
byrandom <- select(byrandom, -all_of(toremove3))
```

## Structure Analysis

```{r}
#| eval: false

ega3 <- bootEGA(
  items,
  EGA.type = "hierEGA",
  model = "glasso",  # BGGM
  algorithm = "louvain",  # walktrap
  type="resampling",
  plot.itemStability=FALSE,
  typicalStructure=TRUE,
  plot.typicalGraph=TRUE,
  iter=500,
  seed=3, ncores = 4)

save(ega3, file="models/ega3.RData")
```

```{r}
load("models/ega3.RData")

# itemStability(ega2)

lower <- make_loadingtable(net.loads(ega3$EGA$lower_order)$std) 
higher <- make_loadingtable(t(net.loads(ega3$EGA$higher_order)$std)) |> 
  select(names(lower)) |> 
  arrange(Item)
higher$Item <- paste0("H", higher$Item)

rbind(higher, lower) |>
  select(-Cluster, -Max) |>
  gt::gt() |>
  gt::tab_row_group(
    label = "Item",
    rows = 1:(nrow(lower) + nrow(higher))
  ) |>
  gt::tab_row_group(
    label = "Higher Order",
    rows = 1:nrow(higher)
  ) |> 
  gt::data_color(columns=-Item,
                 method = "numeric",
                 palette = c("red", "white", "green"),
                 domain = c(-1, 1)) |>
  gt::data_color(columns=Item, fn=\(x) ifelse(x %in% toremove3, "#FFEBEE", "white")) |>
  gt::tab_header(
    title = gt::md("**Item Loadings**"),
    subtitle = "Node centrality"
  )
```

```{r}
graph <- convert2tidygraph(ega3)

tidygraph::tbl_graph(graph$nodes, edges=graph$edges, directed=FALSE) |>
  ggraph::ggraph(layout = 'graphopt') +
  ggraph::geom_edge_arc(aes(edge_color = link, edge_width = abs(link), edge_alpha=as.numeric(type == "real")), strength=0.1) +
  ggraph::geom_edge_arc(aes(edge_alpha=as.numeric(type == "virtual")), edge_width=0.1, strength=0) +
  ggraph::geom_node_point(aes(color=dimension, shape=level), size=10, alpha=0.7) +
  ggraph::geom_node_text(aes(label=name), repel=TRUE) +
  scale_edge_alpha(range = c(0, 0.9), guide="none") +
  scale_shape_manual(values=c("Higher"="square", "Lower"="circle")) +
  theme_void()
```

<!-- ::: {.panel-tabset} -->

<!-- ### Full Sample -->

@@

### Empirical vs. Model Scores

```{r}
EGAnet::net.scores()
```
